G\"{o}del's Ontological Proof of God's Existence (Draft)

### 2013-09-12 Natural Deduction/GodProof-ND.tex
Add ToDo
Fix mistake in Pi_1
Add Annika's KB proof

@@

\begin{prooftree}
\AXC{$ \neg P(\psi)^1$} 
       \AXC{Axiom \ref{A1}} \dashedLine
       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\forall_E$}  
       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp_E$}   
 \BIC{$P(\neg\psi)$}\RightLabel{} 
           	\AXC{$G(x)$} \dottedLine\RightLabel{D\ref{D1}}
		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall_E$}
              	\UIC{$P(\varphi)\imp \varphi(x)$}\RightLabel{$\imp_E$}
           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp_E$}  
            \AXC{$ \psi(x)^2$} \RightLabel{$\imp_E$}  
             \BIC{$\bot $}\RightLabel{$\mathit{RAA}^1$}  
              \UIC{$ P(\psi)$}\RightLabel{$\imp_I^2$}
              \UIC{$ \psi(x)\imp P(\psi)$}
\end{prooftree}

@@

\begin{theorem}
\label{T3}
Necessarily, God exists:
$$
\nec \ex x. G(x)
$$
\end{theorem}
\begin{proof}\hfill
\begin{small}
\begin{prooftree}
\AXC{\textbf{S5}} \dashedLine
\UIC{$ \all \varphi. [\pos \ldots \pos \nec \varphi \biimp \nec \varphi] $}
\UIC{$ \pos \nec \ex x. G(x) \biimp \nec \ex x. G(x) $}
        \AXC{Corollary \ref{C1}} \dashedLine
        \UIC{$\pos \ex x. G(x)$}
                \AXC{Lemma \ref{L1}} \dashedLine
                \UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$} \doubleLine
            \BIC{$\pos \nec \ex x. G(x)$}
    \BIC{$\nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\end{proof}

\section{God exists}

@@

>>>>

ToDo: Do also Scott's and Gödel's proofs.

@@

\begin{prooftree}
\AXC{$ \neg P(\psi)^1$} 
       \AXC{Axiom \ref{A1}} \dashedLine
       \UIC{$\forall \varphi.(\neg P(\varphi)\imp P(\neg\varphi))$}\RightLabel{$\all_E$}  
       \UIC{$\neg P(\psi)\imp P(\neg\psi)$}\RightLabel{$\imp_E$}   
 \BIC{$P(\neg\psi)$}\RightLabel{} 
           	\AXC{$G(x)$} \dottedLine\RightLabel{D\ref{D1}}
		 \UIC{$\forall \varphi .(P(\varphi)\imp \varphi(x))$}\RightLabel{$\forall_E$}
              	\UIC{$P(\neg \psi)\imp \neg \psi(x)$}\RightLabel{$\imp_E$}
           \BIC{$ \neg \psi(x)$}\RightLabel{$\imp_E$}  
            \AXC{$ \psi(x)^2$} \RightLabel{$\imp_E$}  
             \BIC{$\bot $}\RightLabel{$\mathit{RAA}^1$}  
              \UIC{$ P(\psi)$}\RightLabel{$\imp_I^2$}
              \UIC{$ \psi(x)\imp P(\psi)$}
\end{prooftree}

@@

\begin{theorem}
\label{T3}
Necessarily, God exists:

\noindent
\textbf{Theorem:} The existence of a God is necessary:
$$
\nec \ex x. G(x)
$$
\end{theorem}

\noindent
\textbf{Formal proof:}
Let the following derivation be $ \Pi$:

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\nec \ex x. G(x)$} \RightLabel{$\nec$ I}
\UIC{$\nec\nec \ex x. G(x)$}\RightLabel{$\imp$ I, 1}
\UIC{$\nec \ex x. G(x)\imp\nec\nec \ex x. G(x)$} \RightLabel{propositional logic}
\UIC{$\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x)$}\RightLabel{$\nec$ I}
\UIC{$\nec(\neg\nec\nec \ex x. G(x)\imp\neg  \nec \ex x. G(x))$}
\AXC{Axiom K} \dashedLine 
\BIC{$\nec\neg\nec\nec \ex x. G(x)\imp\nec\neg  \nec \ex x. G(x)$} \RightLabel{definition of $\pos$}
 \UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}

\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{1}
\UIC{$\nec\pos\neg\nec \ex x. G(x)$}
\AXC{$ \Pi$} \dashedLine 
\UIC{$\nec\pos\neg\nec \ex x. G(x) \imp\nec\neg\nec \ex x. G(x) $}
 \BIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec$ E, axiom T}
\UIC{$\neg\nec \ex x. G(x)$} 
\AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp$ E}
 \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec$ I}
\UIC{$\nec \neg\ex x. G(x)$}  \RightLabel{definition of $\pos$}
\UIC{$\neg \pos\ex x. G(x)$}

\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $}\dashedLine
\UIC{$\neg \pos\ex x. G(x)$} 
\AXC{ Corollary \ref{C1}} \dashedLine 
\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp$ E}
 \BIC{$ \bot$} \RightLabel{$\imp$ I, 1}
 \UIC{$\neg\nec\pos\neg\nec \ex x. G(x)$}
\AXC{ Axiom B} \dashedLine 
\UIC{$ \neg\nec \ex x. G(x)\imp \nec\pos\neg\nec \ex x. G(x)$} \RightLabel{propositional logic}
\UIC{$ \neg \nec\pos\neg\nec \ex x. G(x)\imp  \nec \ex x. G(x)$} \RightLabel{$\imp$ E}
 \BIC{$ \nec \ex x. G(x)$}
\end{prooftree}
\end{small}

\textbf{Better proof, that directly proves $\ex x. G(x)$:}
\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{1}

 \UIC{$\nec\neg\nec \ex x. G(x)$}\RightLabel{$\nec$ E, axiom T}
\UIC{$\neg\nec \ex x. G(x)$} 
\AXC{$ $} \dashedLine  \RightLabel{Lemma \ref{L1}}
\UIC{$\neg\nec \ex x. G(x) \imp \neg\ex x. G(x)$}\RightLabel{$\imp$ E}
 \BIC{$ \neg\ex x. G(x)$} \RightLabel{$\nec$ I}
\UIC{$\nec \neg\ex x. G(x)$}  \RightLabel{definition of $\pos$}
\UIC{$\neg \pos\ex x. G(x)$} 
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$ $}\dashedLine
\UIC{$\neg \pos\ex x. G(x)$} 
\AXC{ Corollary \ref{C1}} \dashedLine 
\UIC{$ \pos\ex x. G(x)$} \RightLabel{$\imp$ E}
 \BIC{$ \bot$} \RightLabel{$\imp$ I, 1}
 \UIC{$\neg\nec\neg\nec \ex x. G(x)$} \RightLabel{definition of $\pos$}
 \UIC{$\neg\nec\pos \neg \ex x. G(x)$}
\AXC{ Axiom B} \dashedLine 
\UIC{$ \neg \ex x. G(x)\imp\nec\pos \neg \ex x. G(x)$} \RightLabel{propositional logic}
\UIC{$ \neg\nec\pos \neg \ex x. G(x)\imp  \neg  \neg \ex x. G(x)$} \RightLabel{$\imp$ E}
 \BIC{$ \neg  \neg \ex x. G(x)$}\RightLabel{$ \neg  \neg$ E}
 \UIC{$  \ex x. G(x)$}
\end{prooftree}
\end{small}

Note that the last step is classical and we do not prove the existential statement by providing an object for which the statement holds. 
This proof makes section \ref{GodsE} superflous and the use of axiom "M" unnecessary.

The system used contains the $\nec$E-rule with the restriction that we have a $\nec$I below it. This is equivalent to modal sytem K that contains axiom K and necessitation rule N.  We aslo use axiom B ($ A\imp\nec\pos A$). No other modal axioms are needed.

\section{God exists}\label{GodsE}

@@

\section{Proof system equivalent to system K}

We show that the proof system with boxed parts of derivations is equivalent to the system K. The modal system K consists of the axiom K and the necessitation rule N.

\begin{axiom}
\label{AK}
The transitivity axiom K. 
$$
\nec(A\imp B)\imp  (\nec A\imp \nec B)
$$
\end{axiom}

\begin{axiom}
\label{N}
The necessitation rule N.

If $A$ is a theorem, then $\nec A$ is a theorem. 
\end{axiom}

\begin{lemma}
\label{systemK}
The axiom K is derivable in the system. 
\end{lemma}

\begin{small}
\begin{prooftree}
\AXC{$\nec(A\imp B)^2 $}\RightLabel{$\nec$ E}
\UIC{$A\imp B$} 
\AXC{$\nec A ^1 $}\RightLabel{$\nec$ E}
\UIC{$A$} \RightLabel{$\imp$ E}
 \BIC{$ B$} \RightLabel{$\nec$I}
\UIC{$ \nec B$} \RightLabel{$\imp$ I, 1}
 \UIC{$\nec A\imp \nec B$}  \RightLabel{$\imp$ I, 2}
 \UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$}
\end{prooftree}
\end{small}

\begin{lemma}
\label{systemK}
Assuming the axiom K and the necessitation rule N, the open formula $\nec A $ and the existence of a derivation of $ B$ from the open assumption $ A$, then we can derive $\nec B$ without the rules for boxed parts of derivations.
\end{lemma}

\begin{small}
\begin{prooftree}
\AXC{$A^1 $}\dashedLine
\UIC{$ B$} \RightLabel{$\imp$ I, 1}
\UIC{$A\imp B$} \RightLabel{rule N}
\UIC{$\nec(A\imp B)$} 
\AXC{Axiom K}\dashedLine
\UIC{$\nec(A\imp B)\imp  (\nec A\imp \nec B)$} \RightLabel{$\imp$ E}
\BIC{$\nec A\imp \nec B$} 
\AXC{$\nec A ^1 $}\RightLabel{$\imp$ E}
\BIC{$\nec B$}
\end{prooftree}
\end{small}
